---
title: "Green Book Section I.C.1.a - Line point data processing - 2025"
description: "Open-source, reproducible ETL pipeline for processing annual line point transect (LPT) vegetation monitoring data. Raw field data collected on paper data sheets are digitized into Google Sheets, organized by USGS quadrangle with individual parcels in separate worksheets. This automated workflow extracts data from multiple sources (ICWD Google Sheets, LADWP CSV files), applies data quality controls and species code standardization, transforms between long and wide formats, and loads processed datasets into a standardized database schema. The pipeline supports collaborative data management between ICWD and LADWP and ensures reproducibility through version-controlled code and cached intermediate data. Outputs serve as the backend for parameterized reports, interactive dashboards, and decision support tools, enabling transparent vegetation condition assessments."
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    anchor-sections: true
    smooth-scroll: true
    html-math-method: katex
    code-fold: true
    code-summary: "Show code"
date-modified: "2025-12-05"
affiliation: "Inyo County Water Department"
affiliation-url: inyowater.org
categories: [Green Book - Box I.C.1.a.ii, ICWD Annual Report, 2025, vegetation condition]
citation:
  type: report
  container-title: "Annual Report"
  publisher: "Inyo County Water Department"
  issued: "2025-12-05"
  available-date: "2025-12-05"
  url: https://inyo-gov.github.io/vegetation-condition/
google-scholar: true
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
library(DT)
library(googlesheets4)  
library(janitor)
library(dplyr)
library(tidyr)
library(stringr) 
library(readr)

# Load GWalkR if available
if (require(GWalkR, quietly = TRUE)) {
  library(GWalkR)
} else {
  message("GWalkR not available - skipping interactive exploration")
}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(tidyverse.quiet = TRUE)
```

## Overview

This document documents the annual ETL workflow used to convert raw line‑point transect sheets into standardized, analysis‑ready datasets for reporting and QA/QC. Each section below describes the intent of the processing step, with code folded by default to keep the narrative readable.

## Inputs

We start by loading reference data (species codes/traits) used to harmonize species codes and enrich the transect records with taxonomy and functional attributes.

```{r}
# Load species attribute data
species_data <- read.csv("data/species.csv")
```

## Core Transform Functions

These helper functions handle: (1) pivoting from wide transect layout to long format, (2) standardizing types, and (3) joining species attributes and plot IDs. Transect IDs are kept as **character** — regular transects are `"1"`, `"2"`, etc. and suffixed transects retain their original label: `"2_15"`, `"5_20"`, `"3_92"`. The suffix indicates the year the transect was first added (2015, 2020, 1992).

```{r}
# Quad sheets may repeat the species grid below summary rows (QUAD / SPECIES header).
truncate_quad_sheet_repeat_block <- function(df, quad_nm = "QUAD", species_nm = "SPECIES") {
  if (!(quad_nm %in% names(df)) || !(species_nm %in% names(df))) return(df)
  q <- toupper(trimws(as.character(df[[quad_nm]])))
  s <- toupper(trimws(as.character(df[[species_nm]])))
  hdr <- !is.na(q) & !is.na(s) & q == "QUAD" & s == "SPECIES"
  w <- which(hdr)
  if (length(w)) df <- df[seq_len(w[[1L]] - 1L), , drop = FALSE]
  df
}

species_is_numeric_placeholder <- function(species_vals) {
  x <- trimws(as.character(species_vals))
  nzchar(x) & grepl("^-?[0-9]+(?:\\.[0-9]+)?$", x)
}

pivot_longer_icwd <- function(icwd_wide) {
  icwd_wide <- icwd_wide %>%
    mutate(across(starts_with("T"), ~ {
      if (is.list(.)) {
        sapply(., function(x) {
          if (length(x) == 1 && is.numeric(x)) {
            return(as.numeric(x))
          } else {
            return(NA)
          }
        })
      } else {
        return(as.numeric(.))
      }
    }))
  
  measured_columns <- icwd_wide %>%
    select(starts_with("T")) %>%
    select(where(~ any(!is.na(.))))
  
  if (ncol(measured_columns) == 0) {
    return(tibble(Parcel = character(), Transect = character(), Code = character(), Cover = numeric()))
  }
  
  icwd_wide %>%
    pivot_longer(cols = all_of(names(measured_columns)),
                 names_to = "Transect",
                 values_to = "Cover",
                 names_prefix = "T") %>%
    filter(!is.na(Cover)) %>%
    mutate(Transect = as.character(Transect)) %>%
    rename(Code = SPECIES)
}
```

```{r}
# Function to augment data with species attribute information
add_species_agency_plotid <- function(long, cYear, species, entity) {
  long %>%
    mutate(Code = as.character(Code)) %>%  # Ensure Code is a character vector
    left_join(species, by = c("Code" = "Code")) %>%
    mutate(
      source = 'Joint Monitoring 2015-current year',
      source.abr = 'jm',
      Year = cYear,
      Entity = entity,
      plotid = paste(Parcel, Transect, source.abr, sep = '_'),
      plotid.full = paste(Parcel, Transect, source.abr, Entity, sep = '_'),
      Cover = as.numeric(Cover)
    ) %>%
    select(
      Parcel, Code, Transect, Cover, Year, Entity, plotid, Species, CommonName,
      Order, Family, Genus, Lifecycle, Lifeform, Veg_Type, source, source.abr,
      Phreatophyte, plotid.full
    ) %>%
    arrange(Parcel, Transect, Code)
}
```

## Data Sources

The workflow can pull directly from the shared Google Sheets or use a cached intermediate file for faster, reproducible runs. Set `READ_FROM_SHEETS <- TRUE` to refresh from Google Sheets.

```{r}
# Toggle: Set to TRUE to read from Google Sheets, FALSE to use cached intermediary file
READ_FROM_SHEETS <- FALSE

# Define the Google Sheets URLs for 2025
sheets_urls <- list(
  BGP = "https://docs.google.com/spreadsheets/d/1SHMp9g1L2BrGxIK-D6QaJXrzH6iT1iEU1MxtIWsXfqQ/edit?gid=0",
  FSP = "https://docs.google.com/spreadsheets/d/175miDyfxj960js8X2k_cKX-hZOzVGNXDWeOFJcEzzTo/edit?gid=0",
  FSL = "https://docs.google.com/spreadsheets/d/14sXTXykFgRPTERmh9C9GVR5DtPVVx2YLgDKvdCBXtIk/edit?gid=0",
  BLK = "https://docs.google.com/spreadsheets/d/1iuDMhaL5aw_s3wyKj7v60cStu4MpfwgbVRuFhbw6k_E/edit?gid=0",
  IND = "https://docs.google.com/spreadsheets/d/1Hsp0kbB9L85uIVeA21-adBk0xUKnGI441YMBw1huTEM/edit?gid=0",
  MAN = "https://docs.google.com/spreadsheets/d/1sUyBczaBpZKsmLOlcgIufA0ibvWATaOC77y55c_qXxA/edit?gid=0",
  LAW = "https://docs.google.com/spreadsheets/d/10R8eSxxH3aCv55AoBnq7qaHpewxlsr7FQ2N_tz1zPMY/edit?gid=0",
  ABD = "https://docs.google.com/spreadsheets/d/1HACHEcJxZNNxvy8Y1X3LQNaj2l2UOC4kmq1n3Xv5Obc/edit?gid=0",
  PLC = "https://docs.google.com/spreadsheets/d/141gfTxtMvTBfimbPetdTHzvc_0SuOlBZFkn7IKkhjbo/edit?gid=0",
  LNP = "https://docs.google.com/spreadsheets/d/1vIy7R3_DZcV8KdJ7qruTA0S9CDsFpe_uYrljkPJT5Co/edit?gid=0",
  BIS = "https://docs.google.com/spreadsheets/d/1ClvWyz9Wl16d8sGX9-PabdmUMURV1wuYZTIsAhA91As/edit?gid=0",
  TIN = "https://docs.google.com/spreadsheets/d/1-QC__HemASpwMug4Y_kJG1PPSrt3hFHmJ-9X5eWGCKM/edit?gid=0",
  UNW = "https://docs.google.com/spreadsheets/d/1eJRg27m2KCb14rSrk_QjTeI22QBvYTHfGPINL5aklz0/edit?gid=0"
)

# Define path for intermediary file (raw data after reading from sheets, before processing)
intermediary_file <- "output/raw_sheets_data_2025.rds"
```

## Extract (Sheets or Cache)

This step reads all parcel sheets, caches them to `output/raw_sheets_data_2025.rds`, and then proceeds with the same downstream processing regardless of source.

```{r}
# Read data from Google Sheets or load from cached intermediary file
if (READ_FROM_SHEETS) {
  print("Reading data from Google Sheets...")
  
  # Initialize an empty list to store raw dataframes from sheets
  raw_sheets_data <- list()
  
  for (sheet_name in names(sheets_urls)) {
    sheet_url <- sheets_urls[[sheet_name]]
    
    sheets <- sheet_names(sheet_url)  # Get sheet names from Google Sheets
    
    for (sheet in sheets) {
      data <- read_sheet(sheet_url, sheet = sheet)  # Read each sheet
      
      # Print a preview of the data
      print(paste("Reading sheet:", sheet, "from", sheet_name))
      print(paste("Number of rows:", nrow(data)))
      print(paste("Number of columns:", ncol(data)))
      
      # Store raw data with metadata
      if (nrow(data) > 0) {
        raw_sheets_data[[paste(sheet_name, sheet, sep = "_")]] <- list(
          sheet_name = sheet_name,
          sheet = sheet,
          data = data
        )
      }
    }
  }
  
  # Save raw data to intermediary file
  if (!dir.exists("output")) {
    dir.create("output", recursive = TRUE)
  }
  saveRDS(raw_sheets_data, intermediary_file)
  print(paste("Raw data saved to", intermediary_file))
  
} else {
  print("Loading data from cached intermediary file...")
  
  if (file.exists(intermediary_file)) {
    raw_sheets_data <- readRDS(intermediary_file)
    print(paste("Loaded cached data from", intermediary_file))
  } else {
    stop(paste("Intermediary file not found:", intermediary_file, 
               "\nSet READ_FROM_SHEETS <- TRUE to read from Google Sheets first."))
  }
}

# Process the raw data (whether from sheets or cache)
# Initialize an empty list to store processed dataframes
all_data <- list()

for (item in raw_sheets_data) {
  sheet_name <- item$sheet_name
  sheet <- item$sheet
  data <- item$data
  
  print(paste("Processing sheet:", sheet, "from", sheet_name))
  
  # Proceed with filtering and processing only if there are rows
  if (nrow(data) > 0) {
    cleaned_data <- data %>%
      truncate_quad_sheet_repeat_block() %>%
      filter(!grepl("LIVE COVER|LIVE ACRES|STANDARD DEV|COEF OF VAR|ACCURACY|PRN|LIVE,STD,CV,Acc", QUAD, ignore.case = TRUE)) %>%
      select(SPECIES, starts_with("T"), -contains("TOTAL")) %>%
      filter(!species_is_numeric_placeholder(SPECIES))
    
    print(paste("Rows after filtering:", nrow(cleaned_data)))
    
    # Continue processing if there's still data
    if (nrow(cleaned_data) > 0) {
      pivoted_data <- pivot_longer_icwd(cleaned_data) %>%
        mutate(Parcel = sheet)  # Add the sheet name (Parcel) as a new column
      
      summarized_data <- pivoted_data %>%
        group_by(Parcel, Transect, Code) %>%
        summarize(Cover = sum(Cover, na.rm = TRUE), .groups = 'drop')  # Summing cover values
      
      # Fix species code mismatches before joining with species.csv
      # Ensure Code is character (not list) before case_when
      summarized_data <- summarized_data %>%
        mutate(Code = as.character(Code)) %>%
        mutate(Code = case_when(
          Code == "ALOC 2" ~ "ALOC2",
          Code == "CADO" ~ "CADO2",
          Code == "CRRU" ~ "CRRU3",
          Code == "EUOC" ~ "EUOC4",
          Code == "GLLE" ~ "GLLE3",
          Code == "HECU" ~ "HECU3",
          Code == "JUME" ~ "JUME4",
          Code == "MEDIC5" ~ "MEDIC",
          Code == "NIOC" ~ "NIOC2",
          Code == "POFR" ~ "POFR2",
          Code == "SATR" ~ "SATR12",
          Code == "STSP" ~ "STSP3",
          TRUE ~ Code  # Keep original code if no match
        ))
      
      augmented_data <- add_species_agency_plotid(summarized_data, cYear = 2025, species = species_data, entity = "ICWD")
      
      all_data[[length(all_data) + 1]] <- augmented_data
    } else {
      print(paste("No rows left after filtering for sheet:", sheet))
    }
  } else {
    print(paste("No data found in sheet:", sheet))
  }
}

# Combine all the data into one dataframe
final_data <- bind_rows(all_data)

# Check if the final data has rows
if (nrow(final_data) > 0) {
  final_data <- final_data %>%
    arrange(Parcel, Transect, desc(Cover))
  print("Final data preview:")
  print(head(final_data))
} else {
  print("No data to process in the final dataset.")
}
```

## ICWD Processing Output

The extract step produces the ICWD long‑format dataset (`final_data`). This includes species‑code fixes, parcel IDs from sheet names, and per‑transect cover summaries.

```{r}
# Check for duplicates in the final dataset
duplicates <- final_data %>% get_dupes(Parcel, Transect, Code)
if (nrow(duplicates) > 0) {
  print("Duplicates found:")
  print(duplicates)
} else {
  print("No duplicates found in the dataset.")
}
```

## Transect Coverage Summary

Summarize how many transects were read per parcel for the current year to confirm sampling completeness.

```{r}
# Count total transects per year for each parcel
transects_summary <- final_data %>%
  group_by(Parcel, Year) %>%
  summarize(Total_Transects = n_distinct(Transect)) %>%
  arrange(Parcel, Year)

# View the summary
datatable(transects_summary)
```

## Export ICWD Long Format

Save the ICWD long‑format data for downstream joins and QA/QC.

```{r}
# Save the processed data to CSV (long format) — v2 with 400-series transect IDs
write.csv(final_data, "output/icwd_data2025_v2.csv", row.names = FALSE)
print("Data saved to output/icwd_data2025_v2.csv (suffixed transects as 400/500/600-series IDs)")
```

## LADWP Processing

Load LADWP’s long‑format cover file, harmonize codes and transect IDs, and join species attributes to match the ICWD schema.

```{r}
# Process LADWP 2025 data from CSV files
# Read the LADWP Cover data (already in long format)
ladwp_cover_file <- "data/2025_LADWP_LinePoint_Cover.csv"
ladwp_raw <- read.csv(ladwp_cover_file)

print("Processing LADWP 2025 data...")
print(paste("LADWP raw data rows:", nrow(ladwp_raw)))
print(paste("LADWP unique parcels:", length(unique(ladwp_raw$Parcel))))

# Transform LADWP data to match ICWD format
# Rename Species to Code, ensure Transect is character
ladwp_long <- ladwp_raw %>%
  rename(Code = Species) %>%
  mutate(
    Code = as.character(Code),
    Transect = as.character(Transect),
    Cover = as.numeric(Cover)
  ) %>%
  select(Parcel, Transect, Code, Cover) %>%
  filter(!is.na(Cover))

# Fix species code mismatches (same as ICWD)
ladwp_long <- ladwp_long %>%
  mutate(Code = case_when(
    Code == "ALOC 2" ~ "ALOC2",
    Code == "CADO" ~ "CADO2",
    Code == "CRRU" ~ "CRRU3",
    Code == "EUOC" ~ "EUOC4",
    Code == "GLLE" ~ "GLLE3",
    Code == "HECU" ~ "HECU3",
    Code == "JUME" ~ "JUME4",
    Code == "MEDIC5" ~ "MEDIC",
    Code == "NIOC" ~ "NIOC2",
    Code == "POFR" ~ "POFR2",
    Code == "SATR" ~ "SATR12",
    Code == "STSP" ~ "STSP3",
    TRUE ~ Code  # Keep original code if no match
  ))

# Summarize by Parcel, Transect, Code (in case of duplicates)
ladwp_summarized <- ladwp_long %>%
  group_by(Parcel, Transect, Code) %>%
  summarize(Cover = sum(Cover, na.rm = TRUE), .groups = 'drop')

# Augment with species attributes
ladwp_processed <- add_species_agency_plotid(
  ladwp_summarized, 
  cYear = 2025, 
  species = species_data, 
  entity = "LADWP"
)

print("LADWP processed data preview:")
print(head(ladwp_processed, 5))
print(paste("LADWP processed rows:", nrow(ladwp_processed)))
```

## Combine ICWD + LADWP

Bind the two entity datasets into a combined 2025 dataset and export a merged long‑format file.

```{r}
final_data <- final_data %>% mutate(Transect = as.character(Transect))
ladwp_processed <- ladwp_processed %>% mutate(Transect = as.character(Transect))

combined_2025_data <- bind_rows(final_data, ladwp_processed) %>%
  arrange(Parcel, Transect, Code)

print("Combined 2025 data summary:")
print(paste("Total rows:", nrow(combined_2025_data)))
print(paste("ICWD rows:", nrow(final_data)))
print(paste("LADWP rows:", nrow(ladwp_processed)))
print(paste("Unique parcels:", length(unique(combined_2025_data$Parcel))))
print(paste("Unique parcels ICWD:", length(unique(final_data$Parcel))))
print(paste("Unique parcels LADWP:", length(unique(ladwp_processed$Parcel))))

write.csv(combined_2025_data, "output/ICWD_LADWP_merged_lpt_2025_v2.csv", row.names = FALSE)
print("Combined data (v2) saved to output/ICWD_LADWP_merged_lpt_2025_v2.csv")
```

## Master Dataset Update

Append the current year to the prior master file and write a new master CSV.

```{r}
# Create/update master dataset
# Load previous year's master dataset
prev_master_file <- "data/lpt_MASTER_2024.csv"

if (file.exists(prev_master_file)) {
  prev_master <- read.csv(prev_master_file)
  print(paste("Loaded previous master dataset:", nrow(prev_master), "rows"))
  
  prev_master <- prev_master %>% mutate(Transect = as.character(Transect))
  combined_2025_data <- combined_2025_data %>% mutate(Transect = as.character(Transect))
  
  master_2025 <- bind_rows(prev_master, combined_2025_data) %>%
    arrange(Parcel, Year, Transect, Code)
  
  print(paste("Master 2025 dataset:", nrow(master_2025), "rows"))
  print(paste("Years in master:", paste(sort(unique(master_2025$Year)), collapse = ", ")))
  
  write.csv(master_2025, "output/lpt_MASTER_2025_v2.csv", row.names = FALSE)
  print("Master dataset (v2) saved to output/lpt_MASTER_2025_v2.csv")
} else {
  print("Previous master dataset not found. Creating new master from 2025 data only.")
  write.csv(combined_2025_data, "output/lpt_MASTER_2025_v2.csv", row.names = FALSE)
  print("Master dataset (v2) saved to output/lpt_MASTER_2025_v2.csv")
}
```

## Transect Inventory

Summary of all transect IDs per parcel. Suffixed transects (e.g. `2_15`, `3_92`) are kept as-is — the suffix indicates the year the transect was first added.

```{r}
transect_inventory <- final_data %>%
  select(Parcel, Transect, Year) %>%
  distinct() %>%
  mutate(is_suffixed = grepl("_", Transect)) %>%
  arrange(Parcel, Transect)

write.csv(transect_inventory, "output/transect_inventory_2025.csv", row.names = FALSE)
print(paste("Transect inventory:", nrow(transect_inventory), "rows"))

suffixed_only <- transect_inventory %>% filter(is_suffixed)
print(paste("Suffixed transects:", nrow(suffixed_only), "across", n_distinct(suffixed_only$Parcel), "parcels"))
datatable(suffixed_only, caption = "Suffixed transects in 2025 data")
```

## ICWD Wide Export

Pivot ICWD data back to a wide transect format. Suffixed transects (e.g. `T2_15`) are distinct columns, never collapsed into the base transect.

```{r}
sort_transect_cols <- function(tcols) {
  # Sort T-prefixed columns: plain numerics first (T1, T2, ...), then suffixed (T1_15, T2_15, ...)
  stripped <- gsub("^T", "", tcols)
  is_plain <- !grepl("_", stripped)
  plain <- tcols[is_plain]
  suffixed <- tcols[!is_plain]
  plain_sorted <- plain[order(as.numeric(gsub("^T", "", plain)))]
  suffixed_sorted <- suffixed[order(suffixed)]
  c(plain_sorted, suffixed_sorted)
}

wide_format_v2 <- final_data %>%
  select(Parcel, Transect, Code, Cover) %>%
  group_by(Parcel, Code, Transect) %>%
  summarize(Cover = sum(Cover, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    id_cols = c(Parcel, Code),
    names_from = Transect,
    values_from = Cover,
    names_prefix = "T",
    values_fill = NA
  ) %>%
  rename(SPECIES = Code) %>%
  arrange(Parcel, SPECIES)

t_columns <- names(wide_format_v2)[grepl("^T", names(wide_format_v2))]
wide_format_v2 <- wide_format_v2 %>%
  select(Parcel, SPECIES, all_of(sort_transect_cols(t_columns)))

wide_format_output_v2 <- wide_format_v2 %>%
  mutate(across(starts_with("T"), ~ ifelse(is.na(.), "", .)))

write.csv(wide_format_output_v2, "output/icwd_data2025_wide_v2.csv", row.names = FALSE, na = "")
print("Wide format (v2, suffixed transects separated) saved to output/icwd_data2025_wide_v2.csv")

# Production wide input for targets (Code column, resampled transects as separate T* columns)
icwd_wide_production <- wide_format_output_v2 %>%
  rename(Code = SPECIES)
write.csv(icwd_wide_production, "data/lpt_ICWD_2025.csv", row.names = FALSE, na = "")
print("Production ICWD wide input saved to data/lpt_ICWD_2025.csv")
```

## Transect Entity Metadata

Track which parcel+transect records came from ICWD vs LADWP and flag duplicates.

```{r}
# Create metadata table: track which transects belong to which entity
# This helps identify discrepancies and duplications
transect_entity_metadata <- combined_2025_data %>%
  select(Parcel, Transect, Entity, Year) %>%
  distinct() %>%
  arrange(Parcel, Transect, Entity)

# Check for duplicates: same Parcel+Transect in both entities
duplicate_transects <- transect_entity_metadata %>%
  group_by(Parcel, Transect) %>%
  summarize(
    n_entities = n_distinct(Entity),
    entities = paste(sort(unique(Entity)), collapse = ", "),
    .groups = 'drop'
  ) %>%
  filter(n_entities > 1) %>%
  arrange(Parcel, Transect)

print("Transect-Entity Metadata Summary:")
print(paste("Total unique Parcel-Transect combinations:", nrow(transect_entity_metadata)))
print(paste("Parcels with data:", length(unique(transect_entity_metadata$Parcel))))

if (nrow(duplicate_transects) > 0) {
  print(paste("\n⚠️ WARNING: Found", nrow(duplicate_transects), "transects present in multiple entities:"))
  print(duplicate_transects)
} else {
  print("\n✓ No duplicate transects found - each Parcel+Transect appears in only one entity")
}

# Save metadata table
write.csv(transect_entity_metadata, "output/transect_entity_metadata_2025.csv", row.names = FALSE)
print("\nTransect-entity metadata saved to output/transect_entity_metadata_2025.csv")

# Save duplicate report if any exist
if (nrow(duplicate_transects) > 0) {
  write.csv(duplicate_transects, "output/transect_duplicates_2025.csv", row.names = FALSE)
  print("Duplicate transects report saved to output/transect_duplicates_2025.csv")
}
```

## Export LADWP Long Format

Save LADWP’s processed long‑format data for audit and downstream analysis.

```{r}
# Save LADWP processed data separately
write.csv(ladwp_processed, "output/ladwp_data2025.csv", row.names = FALSE)
print("LADWP data saved to output/ladwp_data2025.csv")
```

## LADWP Wide Export

Create a wide‑format LADWP file consistent with the ICWD export.

```{r}
# Export LADWP-only wide format CSV (similar to ICWD)
ladwp_wide_format_v2 <- ladwp_processed %>%
  select(Parcel, Transect, Code, Cover) %>%
  group_by(Parcel, Code, Transect) %>%
  summarize(Cover = sum(Cover, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    id_cols = c(Parcel, Code),
    names_from = Transect,
    values_from = Cover,
    names_prefix = "T",
    values_fill = NA
  ) %>%
  rename(SPECIES = Code) %>%
  arrange(Parcel, SPECIES)

ladwp_t_columns <- names(ladwp_wide_format_v2)[grepl("^T", names(ladwp_wide_format_v2))]
ladwp_wide_format_v2 <- ladwp_wide_format_v2 %>%
  select(Parcel, SPECIES, all_of(sort_transect_cols(ladwp_t_columns)))

ladwp_wide_format_output_v2 <- ladwp_wide_format_v2 %>%
  mutate(across(starts_with("T"), ~ ifelse(is.na(.), "", .)))

write.csv(ladwp_wide_format_output_v2, "output/ladwp_data2025_wide_v2.csv", row.names = FALSE, na = "")
print("LADWP wide format (v2) saved to output/ladwp_data2025_wide_v2.csv")
print(paste("LADWP wide format dimensions:", nrow(ladwp_wide_format_output_v2), "rows,", ncol(ladwp_wide_format_output_v2), "columns"))
```

## Combined Wide Export

Generate a combined ICWD+LADWP wide‑format file for the full 2025 dataset. Suffixed transects appear as 400/500/600-series columns (e.g. T402 for T2_15).

```{r}
combined_wide_format_v2 <- combined_2025_data %>%
  select(Parcel, Transect, Code, Cover) %>%
  group_by(Parcel, Code, Transect) %>%
  summarize(Cover = sum(Cover, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    id_cols = c(Parcel, Code),
    names_from = Transect,
    values_from = Cover,
    names_prefix = "T",
    values_fill = NA
  ) %>%
  rename(SPECIES = Code) %>%
  arrange(Parcel, SPECIES)

combined_t_columns <- names(combined_wide_format_v2)[grepl("^T", names(combined_wide_format_v2))]
combined_wide_format_v2 <- combined_wide_format_v2 %>%
  select(Parcel, SPECIES, all_of(sort_transect_cols(combined_t_columns)))

combined_wide_format_output_v2 <- combined_wide_format_v2 %>%
  mutate(across(starts_with("T"), ~ ifelse(is.na(.), "", .)))

write.csv(combined_wide_format_output_v2, "output/ICWD_LADWP_merged_2025_wide_v2.csv", row.names = FALSE, na = "")
print("Combined ICWD+LADWP wide format (v2) saved to output/ICWD_LADWP_merged_2025_wide_v2.csv")
print(paste("Combined wide format dimensions:", nrow(combined_wide_format_output_v2), "rows,", ncol(combined_wide_format_output_v2), "columns"))
```

## Optional: Interactive Exploration

If available, GWalkR provides an interactive way to explore the ICWD long‑format data.

```{r}
# Optional: Interactive data exploration with GWalkR (if available)
if (require(GWalkR, quietly = TRUE)) {
  GWalkR::gwalkr(final_data, dark = "dark", visConfigFile = 'data/config_gwalk2.json')
} else {
  message("GWalkR not available - skipping interactive exploration")
}
```