---
title: "Green Book Section I.C.1.a — Flagged parcels"
subtitle: "I.C.1.a flagged wellfield profiles — percentile framing, DTW context, and composition"
author: "Inyo County Water Department"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: false
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(targets)
library(tidyverse)
library(knitr)
library(ggplot2)
library(here)
library(patchwork)
library(sf)
library(tmap)
library(tmaptools)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(tidyverse.quiet = TRUE)
```

```{r load-data}
# Same pattern as index.qmd: source report helpers and load targets
source(here("code", "R", "targets_functions.R"))
source(here("code", "R", "report_functions.R"))
source(here("code", "R", "flagged_parcel_context.R"))

targets <- c(
  "cYear",
  "parcels_deltas",
  "deltas_ttest_att",
  "dtw_pfix",
  "rs_pfix",
  "parcels",
  "parcels_shp_ttest",
  "canals_shp",
  "streams_shp",
  "monsites_shp",
  "or_shp",
  "laa_shp",
  "lakes_shp",
  "attributes_pfix"
)
withr::with_dir(rprojroot::find_root("_targets.R"), {
  # rs_pfix is optional — rebuild from parcel_timeseries_app CSV if store object is missing
  req <- setdiff(targets, "rs_pfix")
  tar_load(req)
  tryCatch(tar_load("rs_pfix"), error = function(e)
    message("Optional target rs_pfix not loaded: ", e$message))
})

# Create parcel-pumping linkage once (cached) for parcel profiles
pumping_linkage <- NULL
if (exists("create_parcel_pumping_linkage")) {
  tryCatch({
    linkage_result <- create_parcel_pumping_linkage(
      n_nearest_wells = 10,
      max_distance_m = 10000,
      exempt_max_distance_m = 15000
    )
    pumping_linkage <- linkage_result$parcel_annual_pumping
  }, error = function(e) {
    message("Pumping linkage not available: ", e$message)
  })
} else {
  message("Pumping linkage function not available; skipping pumping summaries.")
}

# App timeseries with Annual_native / Annual_nonnative (same feed as AOP dashboard)
ts_app_path <- here::here("data", "parcel_timeseries_app.csv")
# Prefer current-year export when present
for (cand in c(
  here::here("data", paste0("parcel_timeseries_app_", cYear, ".csv")),
  here::here("data", "parcel_timeseries_app_2026.csv"),
  ts_app_path
)) {
  if (file.exists(cand)) {
    ts_app_path <- cand
    break
  }
}
parcel_ts_app <- if (file.exists(ts_app_path)) {
  readr::read_csv(ts_app_path, show_col_types = FALSE)
} else {
  tibble::tibble()
}

# Percentile framing (precomputed by targets, or on-the-fly from app timeseries)
source(here("code", "R", "measurability_percentiles.R"))
source(here("code", "R", "measurability_ndvi.R"))
pct_summary_path <- here::here("output", "parcel_percentile_summary_flagged.csv")
if (file.exists(pct_summary_path)) {
  parcel_pct_summary <- readr::read_csv(pct_summary_path, show_col_types = FALSE)
} else if (nrow(parcel_ts_app) > 0) {
  parcel_pct_summary <- summarize_parcel_percentiles(
    parcel_metric_percentiles(parcel_ts_app),
    cYear = if (exists("cYear")) cYear else NULL
  )
} else {
  parcel_pct_summary <- tibble::tibble()
}

nmds_index_path <- here::here("output", "nmds_flagged_parcels_index.csv")
nmds_index <- if (file.exists(nmds_index_path)) {
  readr::read_csv(nmds_index_path, show_col_types = FALSE)
} else {
  tibble::tibble(Parcel = character(), nmds_ok = logical(), figure = character())
}

# rs_pfix may be missing from a stale targets store — rebuild NDVI panel feed from app CSV
if (!exists("rs_pfix") || is.null(rs_pfix) || !is.data.frame(rs_pfix) || nrow(rs_pfix) == 0) {
  if (nrow(parcel_ts_app) > 0 && "NDVI_SUR" %in% names(parcel_ts_app)) {
    rs_pfix <- parcel_ts_app %>%
      dplyr::transmute(
        Parcel = .data$Parcel,
        Year = as.integer(.data$Year),
        NDVI_SUR = .data$NDVI_SUR
      ) %>%
      dplyr::filter(is.finite(.data$NDVI_SUR))
    message("rs_pfix rebuilt from ", basename(ts_app_path))
  } else {
    rs_pfix <- tibble::tibble(Parcel = character(), Year = integer(), NDVI_SUR = double())
    message("rs_pfix unavailable; NDVI panels will be empty")
  }
}

# Baseline is parcel-specific. Use the recorded inventory/baseline year rather
# than NominalYear, which can collapse an actual 1987 inventory to 1986.
profile_baseline_year <- function(parcel_id) {
  attr <- attributes_pfix %>%
    dplyr::filter(.data$Parcel == parcel_id) %>%
    dplyr::slice(1)
  candidates <- c(
    if ("Baseline" %in% names(attr)) attr$Baseline else NA,
    if ("bl.Year.tran" %in% names(attr)) attr$`bl.Year.tran` else NA
  )
  years <- suppressWarnings(as.integer(as.character(candidates)))
  years <- years[is.finite(years)]
  if (length(years) > 0) years[[1]] else NA_integer_
}
```

I.C.1.a flagged wellfield parcels (`Cover_sig.counter > 5` or `Grass_sig.counter > 5`). Profiles summarize measurable change vs the 1984–1987 baseline — not cause. Full attributability: [I.C.1.b](https://icwd-attributability.vercel.app).

<details class="section-fold">
<summary><strong>How to interpret a profile</strong></summary>

Wellfield parcels below are I.C.1.a flagged when `Cover_sig.counter > 5` or `Grass_sig.counter > 5` (more than five consecutive years statistically below baseline for total perennial cover or grass). This is a sustained measurable flag under I.C.1.a—not a diagnosis of cause, and not Agreement “significance” under I.C.1.c. Profiles lead with percentile framing (current and baseline years ranked within each parcel’s own time series) and a valley-floor NMDS year path with that parcel’s mapped Green Book TYPE baseline cloud. NMDS distances represent movement through species-composition space (composition geometry / evidence only — not control matching); this informs the evidence picture without serving as the sole I.C.1.a gate. Screening is parcel-by-parcel; questions of causation belong to [I.C.1.b attributability analysis](https://icwd-attributability.vercel.app).

1. **Measurability:** the orange I.C.1.a flag marks a sustained statistical decrease in perennial cover or grass relative to the parcel's baseline.
2. **Support / Not supported (sequential):** Support = decline evidence only — cover↓ and/or grass↓ vs baseline, and/or shrub↑ vs baseline (each metric gated independently; not all required). Sustained counters and percentile rows use the same sign gate. Grass↑, shrub↓, and high grass / still-Type-C proportions go to Not supported (type-stable / contrary). Shown one after the other — not attributability or cause findings.
3. **Hydrologic context:** DTW and precipitation describe water-table and climate conditions. A deeper water table may warrant further attributability analysis; a shallower water table does not by itself establish or rule out causation.
4. **Vegetation and composition:** NDVI is a chlorophyll-related vegetation index (Jul–Sep window noted under the NDVI table). NMDS maps the parcel’s year path on the shared valley-floor plane against its mapped TYPE baseline cloud; percentile summaries and lifeform proportions add complementary evidence. NMDS informs the evidence picture but is not the sole I.C.1.a gate.

The Technical Group may refine methods as field techniques, technology, and analytical understanding develop under Box I.C.1.a.ii.

<details class="section-fold">
<summary><strong>Parcel order</strong> — north→south wellfields (optional)</summary>

Ordered by wellfield, then latitude within wellfield. Full table: [Evaluate change](index.qmd). Map: [Evaluate change — map](index.qmd#map).

</details>

</details>

## Parcel profiles

```{r parcel-profiles, results='asis', fig.height=15, fig.width=11, dpi=120, out.width="100%", fig.align="center"}
# Ensure every plot in this asis loop uses the tall stacked layout
# (opts_chunk$set alone does not resize figures emitted from this same chunk).
knitr::opts_current$set(
  fig.height = 15,
  fig.width = 11,
  dpi = 120,
  out.width = "100%",
  fig.align = "center"
)

# Wellfield order north to south (Control = reference parcels throughout valley)
wellfield_order <- c(
  "Laws", "Control", "Big Pine", "Taboose-Aberdeen",
  "Thibaut Sawmill", "Independence Oak", "Symmes Shepherd", "Bairs Georges"
)

# Only wellfield parcels (exclude Control) where cover OR grass has been below baseline for more than 5 years
# wellfield comes from deltas_ttest_att (join_summaries already joins attributes_pfix); normalize name if CSV has "Wellfield"
profile_parcels <- deltas_ttest_att %>%
  filter(
    Type == "W",
    dplyr::coalesce(Cover_sig.counter, 0) > 5 | dplyr::coalesce(Grass_sig.counter, 0) > 5
  )
if ("Wellfield" %in% names(profile_parcels) && !"wellfield" %in% names(profile_parcels)) {
  profile_parcels <- profile_parcels %>% rename(wellfield = Wellfield)
}
profile_parcels <- profile_parcels %>%
  left_join(
    attributes_pfix %>% select(Parcel, any_of(c("y.cent", "y_cent", "Y.cent"))),
    by = "Parcel"
  ) %>%
  mutate(
    wellfield_group = case_when(
      Type == "C" ~ "Control",
      grepl("Laws", wellfield, ignore.case = TRUE) ~ "Laws",
      grepl("Big Pine", wellfield, ignore.case = TRUE) ~ "Big Pine",
      grepl("Taboose|Aberdeen", wellfield, ignore.case = TRUE) ~ "Taboose-Aberdeen",
      grepl("Thibaut|Sawmill", wellfield, ignore.case = TRUE) ~ "Thibaut Sawmill",
      grepl("Independence|Oak", wellfield, ignore.case = TRUE) ~ "Independence Oak",
      grepl("Symmes|Shepherd", wellfield, ignore.case = TRUE) ~ "Symmes Shepherd",
      grepl("Bairs|Georges", wellfield, ignore.case = TRUE) ~ "Bairs Georges",
      TRUE ~ coalesce(as.character(wellfield), "Other")
    ),
    wellfield_group = factor(wellfield_group, levels = c(wellfield_order, "Other")),
    .y_ord = do.call(dplyr::coalesce, c(as.list(select(., any_of(c("y.cent", "y_cent", "Y.cent")))), list(0)))
  ) %>%
  arrange(wellfield_group, desc(.y_ord), Parcel) %>%
  select(-.y_ord, Parcel, wellfield_group, wellfield, Type, GB_TYPE, everything()) %>%
  head(24)

if (nrow(profile_parcels) == 0) {
  cat("No wellfield parcels meet the I.C.1.a flagged rule (cover or grass statistically below baseline for more than 5 years).")
} else {
  cat("Profiling **", nrow(profile_parcels), "** wellfield parcels on the I.C.1.a flagged list (cover or grass statistically below baseline for more than 5 years).\n\n", sep = "")
}

last_year_if_nonneg <- function(year, delta) {
  yr <- year[!is.na(delta) & delta >= 0]
  if (length(yr) == 0) {
    return(NA_real_)
  }
  max(yr, na.rm = TRUE)
}

last_year_from_counter <- function(counter) {
  if (is.na(counter) || counter <= 0) {
    return(NA_real_)
  }
  cYear - counter
}

# Status badges: DTW OK; shrub ≥80% / grass:shrub cutpoint — Type C only (not Type B)
parcel_gb_type <- function(parcel_id) {
  row <- deltas_ttest_att %>% filter(Parcel == parcel_id) %>% slice(1)
  if (nrow(row) == 0 || !"GB_TYPE" %in% names(row)) return(NA_character_)
  toupper(gsub("[^A-Z/]", "", as.character(row$GB_TYPE[[1]])))
}

type_c_shrub_flag_applies <- function(gb) {
  gb <- toupper(as.character(gb))
  gb %in% c("C", "AC", "A/C", "CA")
}

parcel_status_flags <- function(parcel_id) {
  gb <- parcel_gb_type(parcel_id)
  baseline_year <- profile_baseline_year(parcel_id)
  dtw_current <- dtw_pfix %>%
    filter(Parcel == parcel_id, Year == cYear) %>%
    summarise(DTW = mean(DTW, na.rm = TRUE), .groups = "drop") %>%
    pull(DTW)
  dtw_baseline <- dtw_pfix %>%
    filter(Parcel == parcel_id, Year == baseline_year) %>%
    summarise(DTW = mean(DTW, na.rm = TRUE), .groups = "drop") %>%
    pull(DTW)

  dtw_comparable <- is.finite(dtw_current) && is.finite(dtw_baseline)
  dtw_ok <- dtw_comparable && dtw_current < dtw_baseline
  dtw_deeper <- dtw_comparable && dtw_current >= dtw_baseline

  props <- parcels_deltas %>%
    filter(Parcel == parcel_id, NominalYear == cYear) %>%
    slice(1)
  if (nrow(props) == 0) {
    props <- deltas_ttest_att %>% filter(Parcel == parcel_id) %>% slice(1)
  }
  p_shrub <- NA_real_
  p_grass <- NA_real_
  if (nrow(props) > 0) {
    if ("pShrub" %in% names(props) && is.finite(props$pShrub[[1]])) {
      p_shrub <- props$pShrub[[1]]
    } else if (all(c("Shrub", "Cover") %in% names(props)) &&
               is.finite(props$Cover[[1]]) && props$Cover[[1]] > 0) {
      p_shrub <- props$Shrub[[1]] / props$Cover[[1]]
    }
    if ("pGrass" %in% names(props) && is.finite(props$pGrass[[1]])) {
      p_grass <- props$pGrass[[1]]
    }
  }
  # Prefer perennial stack share when grass+shrub+herb available
  if (nrow(props) > 0 && all(c("pGrass", "pShrub", "pHerb") %in% names(props))) {
    g <- dplyr::coalesce(props$pGrass[[1]], 0)
    s <- dplyr::coalesce(props$pShrub[[1]], 0)
    h <- dplyr::coalesce(props$pHerb[[1]], 0)
    tot <- g + s + h
    if (is.finite(tot) && tot > 0) {
      p_shrub <- s / tot
      p_grass <- g / tot
    }
  }

  shrub_ge_80_raw <- is.finite(p_shrub) && p_shrub >= 0.8
  # Grass:shrub cutpoint ~0.2 grass share (Type C→B reference).
  grass_below_02_raw <- is.finite(p_grass) && p_grass < 0.2
  hist <- parcels_deltas %>%
    filter(Parcel == parcel_id,
           (is.na(baseline_year) | Year > baseline_year),
           is.finite(pShrub)) %>%
    mutate(
      tot = dplyr::coalesce(pGrass, 0) + dplyr::coalesce(pShrub, 0) + dplyr::coalesce(pHerb, 0),
      p_s = ifelse(tot > 0, dplyr::coalesce(pShrub, 0) / tot, pShrub),
      p_g = ifelse(tot > 0, dplyr::coalesce(pGrass, 0) / tot, pGrass)
    )
  shrub_ever_80 <- nrow(hist) > 0 && any(hist$p_s >= 0.8, na.rm = TRUE)
  grass_ever_below_02 <- nrow(hist) > 0 && any(hist$p_g < 0.2, na.rm = TRUE)

  applies <- type_c_shrub_flag_applies(gb)
  dtw_threshold <- dplyr::case_when(
    applies ~ 6,
    identical(gb, "B") ~ 12,
    TRUE ~ NA_real_
  )
  dtw_attention <- isTRUE(dtw_deeper) && is.finite(dtw_threshold) &&
    is.finite(dtw_current) && dtw_current > dtw_threshold
  cover_delta <- if ("Cover.Delta" %in% names(props)) props$Cover.Delta[[1]] else NA_real_
  grass_delta <- if ("Grass.Delta" %in% names(props)) props$Grass.Delta[[1]] else NA_real_
  percentile_for <- function(metric) {
    if (!exists("parcel_pct_summary", inherits = TRUE) ||
        !is.data.frame(parcel_pct_summary)) return(NA_real_)
    hit <- parcel_pct_summary %>%
      dplyr::filter(.data$Parcel == parcel_id, .data$metric == metric) %>%
      dplyr::slice(1)
    if (nrow(hit) == 0) NA_real_ else hit$current_percentile[[1]]
  }
  list(
    dtw_ok = isTRUE(dtw_ok),
    dtw_deeper = isTRUE(dtw_deeper),
    dtw_comparable = isTRUE(dtw_comparable),
    dtw_current = dtw_current,
    dtw_baseline = dtw_baseline,
    dtw_delta = if (isTRUE(dtw_comparable)) dtw_current - dtw_baseline else NA_real_,
    dtw_threshold = dtw_threshold,
    dtw_attention = isTRUE(dtw_attention),
    cover_delta = cover_delta,
    grass_delta = grass_delta,
    cover_percentile = percentile_for("Cover"),
    grass_percentile = percentile_for("Grass"),
    cover_flagged = FALSE,
    gb_type = gb,
    shrub_flag_applies = applies,
    shrub_ge_80 = applies && isTRUE(shrub_ge_80_raw),
    shrub_ge_80_raw = isTRUE(shrub_ge_80_raw),
    shrub_ever_80 = applies && isTRUE(shrub_ever_80),
    grass_cutpoint_flag = applies && isTRUE(grass_below_02_raw),
    grass_cutpoint_raw = isTRUE(grass_below_02_raw),
    grass_ever_below_02 = applies && isTRUE(grass_ever_below_02),
    p_shrub = p_shrub,
    p_grass = p_grass
  )
}

format_parcel_badges <- function(flags) {
  # Three public-facing status cards: change rank, hydrologic context, and composition.
  esc <- function(x) {
    x <- gsub("&", "&amp;", x, fixed = TRUE)
    x <- gsub("<", "&lt;", x, fixed = TRUE)
    x <- gsub(">", "&gt;", x, fixed = TRUE)
    x
  }
  num <- function(x, digits = 1) {
    if (length(x) == 0 || !is.finite(x)) return("—")
    formatC(round(x, digits), format = "f", digits = digits)
  }
  delta <- function(x) {
    if (length(x) == 0 || !is.finite(x)) return(NULL)
    paste0(sprintf("%+.1f", x), " pp")
  }
  ordinal <- function(x) {
    if (length(x) == 0 || !is.finite(x)) return(NULL)
    n <- round(x)
    suffix <- if (n %% 100 %in% 11:13) "th" else switch(as.character(n %% 10),
      "1" = "st", "2" = "nd", "3" = "rd", "th")
    paste0(n, suffix)
  }

  change_bits <- c(
    if (is.finite(flags$cover_delta)) paste0("Cover ", delta(flags$cover_delta)) else NULL,
    if (is.finite(flags$grass_delta)) paste0("Grass ", delta(flags$grass_delta)) else NULL
  )
  pct_bits <- c(
    if (is.finite(flags$cover_percentile)) paste0("Cover ", ordinal(flags$cover_percentile)) else NULL,
    if (is.finite(flags$grass_percentile)) paste0("Grass ", ordinal(flags$grass_percentile)) else NULL
  )
  pct_text <- if (length(pct_bits) == 2 &&
                  identical(ordinal(flags$cover_percentile), ordinal(flags$grass_percentile))) {
    paste0(ordinal(flags$cover_percentile), " percentile")
  } else if (length(pct_bits)) {
    paste(pct_bits, collapse = " · ")
  } else NULL
  change_label <- paste(c(change_bits, pct_text), collapse = " · ")
  if (!nzchar(change_label)) change_label <- "Vegetation change unavailable"
  change_cls <- if (length(change_bits)) "parcel-badge is-bad" else "parcel-badge is-na"

  dtw_direction <- if (is.finite(flags$dtw_delta) && flags$dtw_delta < 0) " (shallower)" else if (is.finite(flags$dtw_delta) && flags$dtw_delta > 0) " (deeper)" else ""
  if (isTRUE(flags$dtw_attention)) {
    dtw_cls <- "parcel-badge is-bad"
    dtw_label <- paste0(
      "DTW ", num(flags$dtw_current, 2), " ft BGS · ",
      sprintf("%+.2f", flags$dtw_delta), " vs baseline", dtw_direction, " · Type ",
      dplyr::coalesce(flags$gb_type, "?"), " >", num(flags$dtw_threshold, 0), " ft"
    )
  } else if (isTRUE(flags$dtw_comparable)) {
    dtw_cls <- "parcel-badge is-good"
    dtw_label <- paste0(
      "DTW ", num(flags$dtw_current, 2), " ft BGS · ",
      sprintf("%+.2f", flags$dtw_delta), " vs baseline", dtw_direction
    )
  } else {
    dtw_cls <- "parcel-badge is-na"
    dtw_label <- "DTW unavailable"
  }

  shrub_pct <- if (is.finite(flags$p_shrub)) paste0(num(100 * flags$p_shrub, 0), "%") else "—"
  grass_pct <- if (is.finite(flags$p_grass)) paste0(num(100 * flags$p_grass, 0), "%") else "—"
  shrub_cls <- if (is.finite(flags$p_shrub)) {
    if (flags$p_shrub > 0.8) "parcel-badge is-bad" else "parcel-badge is-good"
  } else "parcel-badge is-na"
  shrub_label <- paste0("Shrub ", shrub_pct, " · Grass ", grass_pct)
  if (isTRUE(flags$shrub_flag_applies)) shrub_label <- paste0(shrub_label, " · Type C→B >80% screen")

  card <- function(cls, label, aria) sprintf(
    '<span class="%s" role="status" aria-label="%s"><span class="badge-icon" aria-hidden="true">●</span><span class="badge-copy">%s</span></span>',
    cls, esc(aria), esc(label)
  )
  # Status tiles share the same card chrome as Support / Not supported.
  tile <- function(tone, label, aria) {
    tone <- match.arg(tone, c("bad", "good", "na"))
    sprintf(
      paste0(
        '<article class="parcel-card parcel-status-tile is-%s" role="status" aria-label="%s">',
        '<span class="badge-icon" aria-hidden="true">●</span>',
        '<span class="badge-copy">%s</span></article>'
      ),
      tone, esc(aria), esc(label)
    )
  }
  paste0(
    '<div class="parcel-status-row">',
    tile(sub("^parcel-badge is-", "", change_cls), change_label, paste("I.C.1.a", change_label)),
    tile(sub("^parcel-badge is-", "", dtw_cls), dtw_label, dtw_label),
    tile(sub("^parcel-badge is-", "", shrub_cls), shrub_label, shrub_label),
    "</div>\n"
  )
}

escape_html <- function(x) {
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x
}

card_list_html <- function(items) {
  items <- items[!is.na(items) & nzchar(items)]
  if (length(items) == 0) {
    return("<p class=\"parcel-card-empty\">No items yet.</p>\n")
  }
  paste0(
    "<ul class=\"parcel-card-list parcel-next-steps\">\n",
    paste0("  <li><p class=\"next-step-text\">", escape_html(items), "</p></li>", collapse = "\n"),
    "\n</ul>\n"
  )
}

# Two-column cards: Support for measurable change | Next steps
# Optional third card: Support for stable vegetation type (when composition supports type stability)
build_parcel_cards <- function(parcel_id) {
  baseline_year <- profile_baseline_year(parcel_id)
  baseline_label <- if (is.finite(baseline_year)) {
    paste0(baseline_year, " baseline measurement")
  } else {
    "recorded baseline"
  }
  current <- deltas_ttest_att %>%
    filter(Parcel == parcel_id) %>%
    slice(1)

  if (nrow(current) == 0) {
    return("<p>No summary data available for this parcel.</p>\n")
  }

  flags <- parcel_status_flags(parcel_id)
  type_label <- current$GB_TYPE
  wf_label <- current$Type
  wf_name <- if (!is.null(current$wellfield) && !is.na(current$wellfield)) current$wellfield else "Unknown"
  gb <- parcel_gb_type(parcel_id)

  cover_delta <- round(current$Cover.Delta, 1)
  grass_delta <- round(current$Grass.Delta, 1)
  shrub_delta <- round(current$Shrub.Delta, 1)
  pgrass_pct <- if (!is.na(current$pGrass)) round(current$pGrass * 100, 1) else NA_real_
  pshrub_pct <- if (!is.na(current$pShrub)) round(current$pShrub * 100, 1) else NA_real_

  cover_yrs <- dplyr::coalesce(current$Cover_sig.counter, 0)
  grass_yrs <- dplyr::coalesce(current$Grass_sig.counter, 0)

  dtw_current <- dtw_pfix %>%
    filter(Parcel == parcel_id, Year == cYear) %>%
    summarise(DTW = mean(DTW, na.rm = TRUE), .groups = "drop") %>%
    pull(DTW)
  dtw_baseline <- dtw_pfix %>%
    filter(Parcel == parcel_id, Year == baseline_year) %>%
    summarise(DTW = mean(DTW, na.rm = TRUE), .groups = "drop") %>%
    pull(DTW)

  ndvi_now <- NA_real_
  ndvi_base <- NA_real_
  ndvi_row <- rs_pfix %>%
    filter(Parcel == parcel_id, Year %in% c(baseline_year, cYear)) %>%
    group_by(Year) %>%
    summarise(NDVI = mean(NDVI_SUR, na.rm = TRUE), .groups = "drop")
  if (any(ndvi_row$Year == baseline_year)) ndvi_base <- ndvi_row$NDVI[ndvi_row$Year == baseline_year][1]
  if (any(ndvi_row$Year == cYear)) ndvi_now <- ndvi_row$NDVI[ndvi_row$Year == cYear][1]
  # Fallbacks when targets store lacks current year: use recent mean from series
  if (!is.finite(ndvi_now)) {
    ndvi_recent <- rs_pfix %>%
      filter(Parcel == parcel_id, Year >= cYear - 2) %>%
      summarise(NDVI = mean(NDVI_SUR, na.rm = TRUE), .groups = "drop") %>%
      pull(NDVI)
    if (length(ndvi_recent) == 1 && is.finite(ndvi_recent)) ndvi_now <- ndvi_recent
  }

  # Structured evidence cards: each comparison stays aligned by metric rather than prose.
  support_evidence <- list()
  not_supported_evidence <- list()
  next_steps <- c()
  stable_type <- c()
  fmt_num <- function(x, digits = 1) {
    if (length(x) == 0 || !is.finite(x)) return("—")
    formatC(round(x, digits), format = "f", digits = digits)
  }
  fmt_delta <- function(x, digits = 1, suffix = "") {
    if (length(x) == 0 || !is.finite(x)) return("—")
    paste0(sprintf(paste0("%+.", digits, "f"), x), suffix)
  }
  pull_num <- function(name) {
    if (!name %in% names(current)) return(NA_real_)
    val <- suppressWarnings(as.numeric(current[[name]][1]))
    if (length(val) == 0 || !is.finite(val)) NA_real_ else val
  }
  row_simple <- function(label, value, note = NULL) {
    paste0(
      '<div class="evidence-row"><span class="evidence-label">', escape_html(label),
      '</span><span class="evidence-current evidence-wide">', escape_html(value),
      if (!is.null(note)) paste0('<small>', escape_html(note), '</small>') else "",
      '</span></div>'
    )
  }
  # Accounting-style ledger: current | baseline | Δ (Δ is the result column).
  ledger_table <- function(rows, digits = 1, delta_digits = 1, delta_suffix = "",
                           current_suffix = "", baseline_suffix = current_suffix) {
    if (!length(rows)) return("")
    body <- vapply(rows, function(r) {
      paste0(
        '<tr><th scope="row">', escape_html(r$label), '</th>',
        '<td class="num">', escape_html(paste0(fmt_num(r$current, digits), current_suffix)), '</td>',
        '<td class="num">', escape_html(paste0(fmt_num(r$baseline, digits), baseline_suffix)), '</td>',
        '<td class="num delta">', escape_html(fmt_delta(r$delta, delta_digits, delta_suffix)), '</td></tr>'
      )
    }, character(1))
    paste0(
      '<table class="evidence-ledger">',
      '<thead><tr>',
      '<th scope="col" class="ledger-metric"></th>',
      '<th scope="col" class="num">current</th>',
      '<th scope="col" class="num">baseline</th>',
      '<th scope="col" class="num delta">Δ</th>',
      '</tr></thead><tbody>',
      paste(body, collapse = ""),
      '</tbody></table>'
    )
  }
  evidence_card <- function(title, body_html, muted = FALSE) {
    if (!nzchar(paste(body_html, collapse = ""))) return("")
    paste0(
      '<article class="evidence-card', if (muted) ' is-muted' else '', '">',
      '<h4>', escape_html(title), '</h4>', paste(body_html, collapse = ""), '</article>'
    )
  }
  add_split_card <- function(title, support_body, not_body) {
    if (nzchar(paste(support_body, collapse = ""))) {
      support_evidence[[length(support_evidence) + 1]] <<- evidence_card(title, support_body)
    }
    if (nzchar(paste(not_body, collapse = ""))) {
      not_supported_evidence[[length(not_supported_evidence) + 1]] <<- evidence_card(title, not_body, muted = TRUE)
    }
  }
  # DTW still uses row_compare-style one-liner via ledger of one row when needed
  row_compare <- function(label, current_value, baseline_value, delta_value,
                          suffix = "", delta_suffix = suffix, delta_note = "") {
    # kept for DTW single-row; prefer ledger when multi-metric
    paste0(
      '<div class="evidence-row"><span class="evidence-label">', escape_html(label),
      '</span><span class="evidence-current">current ', escape_html(fmt_num(current_value)), suffix,
      '</span><span class="evidence-baseline">baseline ', escape_html(fmt_num(baseline_value)), suffix,
      '</span><span class="evidence-delta">Δ ', escape_html(paste0(fmt_delta(delta_value, 1, delta_suffix), delta_note)),
      '</span></div>'
    )
  }

  # Support sign gate (per metric, independent — not all required):
  #   Cover ↓ (Δ < 0) → Support; Cover ↑/flat → Not supported
  #   Grass ↓ (Δ < 0) → Support; Grass ↑/flat → Not supported (contrary)
  #   Shrub ↑ (Δ > 0) → Support (type-shift / decline composition); Shrub ↓/flat → Not supported
  metric_specs <- tibble::tibble(
    metric = c("Cover", "Grass", "Shrub"),
    current = c(pull_num("Cover"), pull_num("Grass"), pull_num("Shrub")),
    delta = c(cover_delta, grass_delta, shrub_delta),
    # +1 = Support when Δ > 0; -1 = Support when Δ < 0
    support_sign = c(-1L, -1L, 1L)
  ) %>%
    dplyr::mutate(baseline = .data$current - .data$delta)
  change_support_rows <- list()
  change_not_rows <- list()
  for (j in seq_len(nrow(metric_specs))) {
    z <- metric_specs[j, ]
    entry <- list(label = z$metric[[1]], current = z$current[[1]],
                  baseline = z$baseline[[1]], delta = z$delta[[1]])
    d <- z$delta[[1]]
    sgn <- z$support_sign[[1]]
    in_support <- is.finite(d) && ((sgn < 0 && d < 0) || (sgn > 0 && d > 0))
    if (isTRUE(in_support)) {
      change_support_rows[[length(change_support_rows) + 1]] <- entry
    } else {
      change_not_rows[[length(change_not_rows) + 1]] <- entry
    }
  }
  add_split_card(
    "Change vs baseline",
    ledger_table(change_support_rows, digits = 1, delta_digits = 1, delta_suffix = " pp", current_suffix = "%"),
    ledger_table(change_not_rows, digits = 1, delta_digits = 1, delta_suffix = " pp", current_suffix = "%")
  )

  # Same sign gate as Change vs baseline: counter rows Support only when that metric's Δ < 0.
  # yrs > 0 alone is not enough — recovered grass (Δ ≥ 0) belongs under Not supported.
  counter_support <- c()
  counter_not <- c()
  for (nm in c("Cover", "Grass")) {
    yrs <- if (nm == "Cover") cover_yrs else grass_yrs
    delta_m <- if (nm == "Cover") cover_delta else grass_delta
    row <- row_simple(nm, paste0(fmt_num(yrs, 0), " years below baseline"),
                      if (is.finite(yrs) && yrs > 5) "I.C.1.a sustained-change screen" else NULL)
    if (is.finite(yrs) && yrs > 0 && is.finite(delta_m) && delta_m < 0) {
      counter_support <- c(counter_support, row)
    } else {
      counter_not <- c(counter_not, row)
    }
  }
  add_split_card("Sustained counters", counter_support, counter_not)

  pct_rows <- if (exists("parcel_pct_summary") && is.data.frame(parcel_pct_summary)) {
    parcel_pct_summary %>% dplyr::filter(.data$Parcel == parcel_id)
  } else tibble::tibble()
  # Percentile rows follow the same metric Δ sign gate as Change vs baseline
  # (Cover↓ / Grass↓ / Shrub↑ → Support; opposite signs → Not supported).
  # Do not use cur_p < 50: a mid-series percentile can be <50 while Δ vs baseline is positive
  # (e.g. FSL054 Grass CURRENT 32 vs BASELINE 5, Δ +27) — that is contrary / Not supported.
  pct_support_rows <- list()
  pct_not_rows <- list()
  for (met in c("Cover", "Grass", "Shrub")) {
    pr <- pct_rows %>% dplyr::filter(.data$metric == met) %>% dplyr::slice(1)
    if (!nrow(pr)) next
    cur_p <- pr$current_percentile[[1]]
    base_p <- pr$baseline_percentile_mean[[1]]
    delta_p <- if (is.finite(cur_p) && is.finite(base_p)) cur_p - base_p else NA_real_
    entry <- list(label = met, current = cur_p, baseline = base_p, delta = delta_p)
    abs_delta <- if (met == "Cover") cover_delta else if (met == "Grass") grass_delta else shrub_delta
    # Cover/Grass: Support on decline (Δ < 0). Shrub: Support on increase (Δ > 0).
    in_support <- if (met == "Shrub") {
      is.finite(abs_delta) && abs_delta > 0
    } else {
      is.finite(abs_delta) && abs_delta < 0
    }
    if (isTRUE(in_support)) {
      pct_support_rows[[length(pct_support_rows) + 1]] <- entry
    } else {
      pct_not_rows[[length(pct_not_rows) + 1]] <- entry
    }
  }
  add_split_card(
    "Percentile rank",
    ledger_table(pct_support_rows, digits = 0, delta_digits = 0, delta_suffix = " pct pts"),
    ledger_table(pct_not_rows, digits = 0, delta_digits = 0, delta_suffix = " pct pts")
  )

  swatch <- function(hex) paste0('<span class="lifeform-swatch" style="background:', hex, '" aria-hidden="true"></span>')
  proportion_rows <- c(
    if (is.finite(pgrass_pct)) paste0(
      '<div class="evidence-row"><span class="evidence-label">', swatch("#1B5E20"),
      'Grass</span><span class="evidence-current evidence-wide">',
      escape_html(paste0(fmt_num(pgrass_pct, 0), "% of perennial cover")),
      '</span></div>'
    ) else NULL,
    if (is.finite(pshrub_pct)) paste0(
      '<div class="evidence-row"><span class="evidence-label">', swatch("#8D6E63"),
      'Shrub</span><span class="evidence-current evidence-wide">',
      escape_html(paste0(fmt_num(pshrub_pct, 0), "% of perennial cover")),
      '</span></div>'
    ) else NULL
  )
  if (length(proportion_rows)) {
    # High grass / still-Type-C (meadow-like) → Not supported (type-stable).
    # Type C shrub≥80% drift alarm → Support (composition change, not type-stable).
    type_stable_props <- (is.finite(pgrass_pct) && pgrass_pct >= 60) ||
      (isTRUE(flags$shrub_flag_applies) && !isTRUE(flags$shrub_ge_80)) ||
      (!isTRUE(flags$shrub_flag_applies) && identical(gb, "C"))
    type_drift_props <- isTRUE(flags$shrub_flag_applies) && isTRUE(flags$shrub_ge_80)
    if (type_drift_props) {
      support_evidence[[length(support_evidence) + 1]] <- evidence_card(
        "Perennial proportions (type drift)", proportion_rows
      )
    } else if (type_stable_props || length(proportion_rows)) {
      not_supported_evidence[[length(not_supported_evidence) + 1]] <- evidence_card(
        "Perennial proportions (type-stable)", proportion_rows, muted = TRUE
      )
    }
  } else {
    not_supported_evidence[[length(not_supported_evidence) + 1]] <- evidence_card(
      "Perennial proportions", row_simple("Composition", "Unavailable"), muted = TRUE
    )
  }

  dtw_delta <- if (is.finite(dtw_current) && is.finite(dtw_baseline)) dtw_current - dtw_baseline else NA_real_
  dtw_row <- if (is.finite(dtw_current) && is.finite(dtw_baseline)) {
    row_compare(
      "DTW (ft BGS)", dtw_current, dtw_baseline, dtw_delta, "", " ft",
      if (dtw_delta < 0) " (shallower)" else if (dtw_delta > 0) " (deeper)" else ""
    )
  } else row_simple("DTW", "Unavailable")
  # DTW is hydrologic context, not decline evidence → Not supported column.
  not_supported_evidence[[length(not_supported_evidence) + 1]] <- evidence_card(
    "Hydrologic context", dtw_row, muted = TRUE
  )

  # NDVI: compact table + residual screen (unexpected baseline offset → investigate baseline validity)
  if (is.finite(ndvi_now) || is.finite(ndvi_base)) {
    ndvi_year_label <- if (any(ndvi_row$Year == cYear)) as.character(cYear) else "most recent"
    # Build Cover+NDVI series for residual screen (app CSV / parcels_deltas)
    ndvi_ts_join <- NULL
    if (exists("parcels_deltas") && is.data.frame(parcels_deltas)) {
      ndvi_ts_join <- parcels_deltas %>%
        dplyr::filter(.data$Parcel == parcel_id) %>%
        dplyr::transmute(
          Year = as.integer(.data$NominalYear),
          Cover = as.numeric(.data$Cover),
          NDVI_SUR = NA_real_
        )
      if (exists("rs_pfix") && is.data.frame(rs_pfix) && nrow(rs_pfix)) {
        ndvi_ts_join <- ndvi_ts_join %>%
          dplyr::left_join(
            rs_pfix %>%
              dplyr::filter(.data$Parcel == parcel_id) %>%
              dplyr::group_by(.data$Year) %>%
              dplyr::summarise(NDVI_SUR = mean(.data$NDVI_SUR, na.rm = TRUE), .groups = "drop"),
            by = "Year"
          )
      }
    }
    if (exists("parcel_ts_app") && is.data.frame(parcel_ts_app) &&
        all(c("Parcel", "Cover", "NDVI_SUR") %in% names(parcel_ts_app))) {
      year_col <- if ("Year" %in% names(parcel_ts_app)) "Year" else if ("NominalYear" %in% names(parcel_ts_app)) "NominalYear" else NULL
      if (!is.null(year_col)) {
        ndvi_ts_join <- parcel_ts_app %>%
          dplyr::filter(.data$Parcel == parcel_id) %>%
          dplyr::transmute(
            Year = as.integer(.data[[year_col]]),
            Cover = as.numeric(.data$Cover),
            NDVI_SUR = as.numeric(.data$NDVI_SUR)
          )
      }
    }
    ndvi_screen <- if (!is.null(ndvi_ts_join) && nrow(ndvi_ts_join)) {
      ndvi_cover_residual_screen(ndvi_ts_join)
    } else {
      tibble::tibble(ndvi_evidence_problematic = FALSE, baseline_resid_pp = NA_real_,
                     baseline_validity_note = NA_character_)
    }
    ndvi_flag_html <- if (isTRUE(ndvi_screen$ndvi_evidence_problematic[[1]])) {
      paste0(
        '<p class="evidence-flag is-problem">',
        '<strong>NDVI evidence problematic.</strong> ',
        escape_html(ndvi_screen$baseline_validity_note[[1]]),
        '</p>'
      )
    } else if (is.finite(ndvi_screen$baseline_resid_pp[[1]])) {
      paste0(
        '<p class="evidence-footnote">Baseline Cover residual vs NDVI–cover fit: ',
        escape_html(sprintf("%+.1f pp", ndvi_screen$baseline_resid_pp[[1]])),
        ' (placeholder screen ±', NDVI_BASELINE_RESID_PP_THRESHOLD, ' pp).</p>'
      )
    } else ""
    ndvi_table <- paste0(
      '<table class="evidence-mini-table"><thead><tr><th scope="col">Period</th><th scope="col">NDVI</th></tr></thead><tbody>',
      '<tr><td>Most recent (', escape_html(ndvi_year_label), ')</td><td>',
      escape_html(if (is.finite(ndvi_now)) formatC(ndvi_now, digits = 3, format = "f") else "—"),
      '</td></tr>',
      '<tr><td>Baseline (', escape_html(baseline_year), ')</td><td>',
      escape_html(if (is.finite(ndvi_base)) formatC(ndvi_base, digits = 3, format = "f") else "—"),
      '</td></tr>',
      if (is.finite(ndvi_now) && is.finite(ndvi_base))
        paste0(
          '<tr><td>Δ (recent − baseline)</td><td>',
          escape_html(sprintf("%+.3f", ndvi_now - ndvi_base)),
          '</td></tr>'
        )
      else "",
      '</tbody></table>',
      ndvi_flag_html,
      '<p class="evidence-footnote">[Jul 15–Sep 15 aggregate; roughly two months after the June–late-July LPT sampling window.]</p>'
    )
    # Inline 1:1 / residual diagnostic plot when present
    ndvi_1to1 <- here::here("output", "ndvi_cover_baseline", paste0(parcel_id, "_ndvi_cover_1to1.png"))
    ndvi_resid_png <- here::here("output", "figures", "ndvi_residuals", paste0(parcel_id, "_ndvi_residual.png"))
    diag_png <- if (file.exists(ndvi_1to1)) ndvi_1to1 else if (file.exists(ndvi_resid_png)) ndvi_resid_png else NA_character_
    if (!is.na(diag_png)) {
      ndvi_table <- paste0(
        ndvi_table,
        '<figure class="ndvi-inline-diag"><img src="', knitr::image_uri(diag_png),
        '" alt="', escape_html(parcel_id), ' NDVI-cover diagnostic"/>',
        '<figcaption>NDVI↔cover 1:1 diagnostic with separate OLS for pre-2015 (re-randomized) vs 2015+ (permanent); dashed = standardized 1:1. ',
        'Unexpected baseline offset → investigate baseline validity.</figcaption></figure>'
      )
    }
    # NDVI is vegetation-index / baseline-validity context, not decline evidence → Not supported.
    ndvi_title <- if (isTRUE(ndvi_screen$ndvi_evidence_problematic[[1]])) {
      "NDVI vegetation index (baseline validity)"
    } else {
      "NDVI vegetation index"
    }
    not_supported_evidence[[length(not_supported_evidence) + 1]] <- evidence_card(
      ndvi_title, ndvi_table, muted = TRUE
    )
  }

  # Grass proportion stability → stable veg type (not an alarm)
  pgrass_hist <- parcels_deltas %>%
    filter(Parcel == parcel_id,
           (is.na(baseline_year) | Year > baseline_year),
           is.finite(pGrass)) %>%
    arrange(Year)
  if (nrow(pgrass_hist) >= 3) {
    mn <- round(100 * mean(pgrass_hist$pGrass, na.rm = TRUE), 0)
    lo <- round(100 * min(pgrass_hist$pGrass, na.rm = TRUE), 0)
    hi <- round(100 * max(pgrass_hist$pGrass, na.rm = TRUE), 0)
    cur_g <- if (is.finite(flags$p_grass)) round(100 * flags$p_grass, 0) else pgrass_pct
    if (is.finite(cur_g) && cur_g >= 60) {
      stable_type <- c(
        stable_type,
        paste0(
          "Grass still ~", cur_g, "% of perennial cover (reinventory mean ~", mn,
          "%; range ", lo, "–", hi, "%). Meadow/grass dominance supports vegetation-type stability — not a shrub-encroachment alarm."
        )
      )
    }
    if (!isTRUE(flags$shrub_flag_applies) && isTRUE(flags$shrub_ge_80_raw)) {
      stable_type <- c(
        stable_type,
        paste0(
          "Mapped Type ", dplyr::coalesce(gb, type_label),
          ": high shrub share is consistent with type; the Type C→B shrub≥80% / grass:shrub≈0.2 cutpoint flag does not apply."
        )
      )
    } else if (isTRUE(flags$shrub_flag_applies) && !isTRUE(flags$shrub_ge_80) && is.finite(pshrub_pct) && pshrub_pct < 80) {
      stable_type <- c(
        stable_type,
        "Shrub share below the Type C→B ≥80% reference — high-shrub alarm does not apply at present."
      )
    }
  }

  # Type B if-then framing
  if (identical(gb, "B")) {
    next_steps <- c(
      next_steps,
      paste0(
        "Mapped Type B (shrubland) context: a sustained cover decline paired with a deeper-than-expected water table ",
        "may warrant I.C.1.b attributability analysis. If the water table is not deeper than expected for Type B, ",
        "interpret cover change within the normal range of mapped Type B shrublands rather than treating it as a meadow-type (C) alarm. ",
        "The Type C→B shrub≥80% / grass:shrub cutpoint does not apply while the parcel remains mapped Type B."
      )
    )
  }

  # FSL054 special consideration
  if (identical(parcel_id, "FSL054")) {
    next_steps <- c(
      next_steps,
      paste0(
        "Special consideration — riparian history: this stand was historically riparian scrub / coyote willow (Type D character). ",
        "After pumping-related die-off, fire, and river-level / irrigation practices that no longer support Type D, ",
        "the vegetation is now more akin to Type C meadow. Treat mapped type vs present composition as an interpretive context, ",
        "not a simple C→B shrub-encroachment story."
      ),
      paste0(
        "Composition differs from the mapped type because of that riparian history; use those differences for interpretation ",
        "and future management analysis, not as an automatic Type C alarm."
      ),
      paste0(
        "Naming note: FSL* parcel codes here are Laws wellfield parcels along the Owens River mainstem — not Fish Slough."
      )
    )
  }

  # Next steps / attributability follow-ups
  if (is.finite(dtw_current) && is.finite(dtw_baseline)) {
    if (dtw_current < dtw_baseline) {
      next_steps <- c(
        next_steps,
        paste0(
          paste0("Depth-to-water (DTW) is currently shallower than the ", baseline_label, " on this parcel. "),
          paste0("In plain terms, the water table is closer to the surface than it was at the ", baseline_label, ". "),
          "The Green Book “adequacy-for-type” hydro screen asks whether depth is too deep for the mapped vegetation type; ",
          "a shallower reading does not, by itself, push that screen. ",
          "It also does not prove or disprove pumping as a cause — it only means hydrology alone is not the red flag here. ",
          "Still check composition change, baseline quality (including NDVI↔cover residuals), and other I.C.1.b factors before causal language."
        )
      )
    } else {
      next_steps <- c(
        next_steps,
        paste0(
          paste0("Depth-to-water (DTW) is currently deeper than the ", baseline_label, " (water table farther below ground). "),
          "If that deeper condition is sustained and looks incompatible with mapped Type ", type_label,
          ", prioritize I.C.1.b water-table / soil-water factors in attributability analysis — ",
          "together with rainfall, land use, and other stressors. Deeper DTW is context for causation review, not a final “but for” finding by itself."
        )
      )
    }
  } else {
    next_steps <- c(
      next_steps,
      paste0(
        "DTW vs baseline is unavailable for this parcel in the current tables. ",
        "Attributability (I.C.1.b) needs hydrologic context before “but for pumping / surface-water practice” language; ",
        "pull well records or revisit the DTW join before closing a hydro screen."
      )
    )
  }

  if (identical(parcel_id, "FSL044")) {
    next_steps <- c(
      next_steps,
      paste0(
        "NDVI↔cover diagnostic (plot in the margin / NDVI card): 1987 baseline cover sits well above the parcel’s NDVI–cover fit ",
        "(large positive residual). That pattern is a flag to investigate baseline validity (possible inflated baseline), ",
        "not a causal conclusion about pumping."
      ),
      paste0(
        "Pre-2015 vs 2015+ permanent-transect network eras differ in sampling design; ",
        "do not over-interpret single-year NDVI/cover mismatches across that break."
      ),
      paste0(
        "Meadow-dominant composition and currently shallow DTW are important context for any later attributability discussion on this parcel."
      )
    )
  }

  if (identical(gb, "B")) {
    next_steps <- c(
      next_steps,
      paste0(
        "Keep the grass:shrub ratio visible in the timeseries for composition context only. ",
        "Do not flag or badge on the Type C→B shrub≥80% / ~0.2 grass-share cutpoint while the parcel is mapped Type B."
      )
    )
  }

  if (!identical(parcel_id, "FSL054") && cover_yrs > 5) {
    next_steps <- c(
      next_steps,
      paste0(
        "Measurability under Box I.C.1.a.ii is already established for this parcel (sustained cover and/or grass below baseline). ",
        "Causation is a separate step: I.C.1.b attributability requires control comparisons, rainfall, land use, and other stressors ",
        "before any “but for pumping / surface-water practice” conclusion."
      )
    )
  }



  # Sequential screening: Support, then Not supported (no top status tiles; no side-by-side columns).
  html <- paste0(
    '<div class="parcel-screening parcel-screening-stack">\n',
    '<p class="parcel-cards-intro">',
    '<strong>Support</strong> = <em>decline evidence only</em> (cover↓ / grass↓ / shrub↑ vs baseline; sustained counters and percentile rows on the same sign gate). ',
    '<strong>Not supported</strong> = type-stable / contrary signals (grass↑ / shrub↓ / cover↑; high grass / still-Type-C proportions; DTW &amp; NDVI context). ',
    'Not attributability or cause (see I.C.1.b).',
    '</p>\n',
    '<div class="parcel-cards parcel-cards-stack">\n',
    '<section class="parcel-card parcel-card-support">\n',
    '<p class="parcel-card-title">Support <span class="parcel-card-sub">(decline evidence)</span></p>\n',
    if (length(support_evidence)) paste(support_evidence, collapse = "") else '<p class="parcel-card-empty">No decline evidence identified.</p>',
    '</section>\n',
    '<section class="parcel-card parcel-card-not-supported">\n',
    '<p class="parcel-card-title">Not supported <span class="parcel-card-sub">(type-stable / contrary)</span></p>\n',
    if (length(not_supported_evidence)) paste(not_supported_evidence, collapse = "") else '<p class="parcel-card-empty">No type-stable / contrary signal identified.</p>',
    '</section>\n',
    '<section class="parcel-card parcel-card-next">\n',
    '<p class="parcel-card-title">Next steps</p>\n',
    card_list_html(unique(next_steps)),
    '</section>\n'
  )
  if (length(stable_type) > 0) {
    html <- paste0(
      html,
      '<section class="parcel-card parcel-card-stable">\n',
      '<p class="parcel-card-title">Type-stable notes <span class="parcel-card-sub">(also under Not supported)</span></p>\n',
      card_list_html(unique(stable_type)),
      '</section>\n'
    )
  }
  paste0(html, "</div>\n</div>\n")
}

# Back-compat alias used nowhere after this patch
build_parcel_notes <- function(parcel_id) {
  # retained for any external callers; prefer build_parcel_cards
  character(0)
}

for (i in seq_len(nrow(profile_parcels))) {
  pid <- profile_parcels$Parcel[i]
  baseline_year <- profile_baseline_year(pid)
  wf_group <- as.character(profile_parcels$wellfield_group[i])
  wf_slug <- sub("-+$", "", gsub("[^a-z0-9]+", "-", tolower(wf_group)))
  anchor_id <- paste0(wf_slug, "--parcel-", tolower(pid))
  cat("\n\n### ", wf_group, " — Parcel ", pid, " {#", anchor_id, "}\n\n", sep = "")

  # Full-width map band first (when assets exist), then cards + figs below — never trap
  # the map beside NMDS/residual margin or a skinny left column.
  if (exists("emit_flagged_parcel_context", mode = "function")) {
    tryCatch(emit_flagged_parcel_context(pid), error = function(e)
      message("flagged context skip for ", pid, ": ", e$message))
  }

  cat('<div class="parcel-profile-layout">\n')
  cat('<div class="parcel-profile-main">\n')
  cat(build_parcel_cards(pid))
  cat('<div class="parcel-timeseries-block">\n')

  ts_raw <- parcels_deltas %>%
    filter(Parcel == pid) %>%
    select(Year, Cover, Grass, Shrub, Herb, pGrass, pShrub, pHerb) %>%
    arrange(Year)

  year_min <- 1985
  year_max <- cYear

  current_vals <- parcels_deltas %>%
    filter(Parcel == pid, Year == cYear) %>%
    select(Cover, Grass, Shrub, Herb, pGrass, pShrub, pHerb) %>%
    slice(1)

  if (nrow(current_vals) == 0) {
    current_vals <- deltas_ttest_att %>%
      filter(Parcel == pid) %>%
      slice(1) %>%
      select(any_of(c("Cover", "Grass", "Shrub", "Herb", "pGrass", "pShrub", "pHerb")))
  }

  fill_pGrass <- if (nrow(current_vals) > 0 && "pGrass" %in% names(current_vals)) current_vals$pGrass[[1]] else NA_real_
  fill_pShrub <- if (nrow(current_vals) > 0 && "pShrub" %in% names(current_vals)) current_vals$pShrub[[1]] else NA_real_
  fill_pHerb <- if (nrow(current_vals) > 0 && "pHerb" %in% names(current_vals)) current_vals$pHerb[[1]] else NA_real_

  ts <- ts_raw %>%
    complete(Year = year_min:year_max) %>%
    mutate(
      pGrass = ifelse(Year == cYear & is.na(pGrass), fill_pGrass, pGrass),
      pShrub = ifelse(Year == cYear & is.na(pShrub), fill_pShrub, pShrub),
      pHerb = ifelse(Year == cYear & is.na(pHerb), fill_pHerb, pHerb),
      # Fill proportions from absolute cover when needed
      pGrass = ifelse(is.na(pGrass) & is.finite(Cover) & Cover > 0 & is.finite(Grass), Grass / Cover, pGrass),
      pShrub = ifelse(is.na(pShrub) & is.finite(Cover) & Cover > 0 & is.finite(Shrub), Shrub / Cover, pShrub),
      pHerb = ifelse(is.na(pHerb) & is.finite(Cover) & Cover > 0 & is.finite(Herb), Herb / Cover, pHerb),
      pGrass = round(pGrass, 3),
      pShrub = round(pShrub, 3),
      pHerb = round(pHerb, 3)
    )

  # Prefer current year on the axis; drop prior 5-yr tick if it would collide (e.g. 2025 vs 2026)
  year_breaks <- seq(year_min, year_max, by = 5)
  if (!(year_max %in% year_breaks)) {
    if (length(year_breaks) && (year_max - max(year_breaks)) < 3) {
      year_breaks <- year_breaks[year_breaks < max(year_breaks)]
    }
    year_breaks <- c(year_breaks, year_max)
  }
  # Room for y titles + tick labels so stacked panels don't collide
  plot_margin <- margin(12, 20, 14, 28)
  panel_theme <- theme_minimal(base_size = 12) +
    theme(
      plot.margin = plot_margin,
      axis.title.y = element_text(size = 11, margin = margin(r = 16)),
      axis.text.y = element_text(size = 9),
      axis.text.x = element_text(size = 9),
      legend.text = element_text(size = 9),
      legend.title = element_text(size = 9),
      plot.caption = element_text(hjust = 0, size = 8.5, color = "gray40")
    )
  x_limits <- c(year_min - 0.5, year_max + 0.75)

  # Shared lifeform palette (stack + line + proportion cards)
  # Grass kept dark green; herb purple; shrub brown; annuals light/dark pink
  col_grass <- "#1B5E20"
  col_herb <- "#7B1FA2"
  col_shrub <- "#8D6E63"
  col_cover <- "#263238"
  col_ann_nat <- "#F8BBD0"
  col_ann_non <- "#C2185B"

  # Absolute cover — the series that drives Cover/Grass measurability flags
  baseline_cover <- ts %>%
    filter(Year == baseline_year) %>%
    summarise(
      Cover = mean(Cover, na.rm = TRUE),
      Grass = mean(Grass, na.rm = TRUE),
      .groups = "drop"
    )

  ts_cover <- ts %>%
    select(Year, Cover, Grass, Shrub) %>%
    pivot_longer(cols = c(Cover, Grass, Shrub), names_to = "metric", values_to = "value") %>%
    mutate(
      metric = factor(
        metric,
        levels = c("Cover", "Grass", "Shrub"),
        labels = c("Total cover", "Grass", "Shrub")
      )
    )

  cover_end <- ts_cover %>%
    dplyr::filter(is.finite(.data$value)) %>%
    dplyr::group_by(.data$metric) %>%
    dplyr::slice_max(order_by = .data$Year, n = 1, with_ties = FALSE) %>%
    dplyr::ungroup()

  # Line/points only where Cover/Grass/Shrub are observed; NA years (no survey)
  # leave gaps — do not connect across missing inventory years.
  p_cover <- ggplot(ts_cover, aes(x = Year, y = value, color = metric)) +
    geom_line(linewidth = 0.7, na.rm = FALSE) +
    geom_point(size = 1.8, na.rm = TRUE) +
    geom_text(
      data = cover_end,
      aes(label = metric),
      hjust = -0.08, vjust = 0.5, size = 2.7, show.legend = FALSE, check_overlap = TRUE
    ) +
    scale_color_manual(
      values = c("Total cover" = col_cover, Grass = col_grass, Shrub = col_shrub)
    ) +
    labs(x = NULL, y = "Cover (%)", color = NULL) +
    scale_x_continuous(breaks = year_breaks, expand = expansion(mult = c(0, 0))) +
    coord_cartesian(xlim = c(year_min - 0.5, year_max + 3.2), clip = "off") +
    panel_theme +
    theme(legend.position = "bottom") +
    guides(color = guide_legend(nrow = 1, override.aes = list(linewidth = 1.2, size = 2.5)))

  if (is.finite(baseline_cover$Cover[1])) {
    p_cover <- p_cover +
      geom_hline(
        yintercept = baseline_cover$Cover[1],
        linetype = "dashed", color = col_cover, linewidth = 0.35
      ) +
      annotate(
        "text", x = year_min, y = baseline_cover$Cover[1],
        label = paste0("Cover ", baseline_year, " baseline"), hjust = 0, vjust = -0.4, size = 2.8, color = col_cover
      )
  }
  if (is.finite(baseline_cover$Grass[1]) &&
      (!is.finite(baseline_cover$Cover[1]) ||
         abs(baseline_cover$Grass[1] - baseline_cover$Cover[1]) > 1)) {
    p_cover <- p_cover +
      geom_hline(
        yintercept = baseline_cover$Grass[1],
        linetype = "dashed", color = col_grass, linewidth = 0.35
      ) +
      annotate(
        "text", x = year_min, y = baseline_cover$Grass[1],
        label = paste0("Grass ", baseline_year, " baseline"), hjust = 0, vjust = 1.2, size = 2.8, color = col_grass
      )
  }

  # Cover stack with annual native / non-native (AOP dashboard colors & series)
  # Only survey/inventory years (finite Cover). Do NOT expand the year grid and
  # fill with 0 — missing LPT years must stay empty (no bar, no total-cover dot).
  p_annual_stack <- NULL
  if (nrow(parcel_ts_app) > 0 && pid %in% parcel_ts_app$Parcel) {
    stack_ts <- parcel_ts_app %>%
      dplyr::filter(.data$Parcel == pid, .data$Year >= year_min, .data$Year <= year_max) %>%
      dplyr::filter(is.finite(.data$Cover)) %>%
      dplyr::mutate(
        # Within an observed survey year, absent lifeform components are true 0
        Grass = dplyr::coalesce(.data$Grass, 0),
        Shrub = dplyr::coalesce(.data$Shrub, 0),
        Cover = .data$Cover,
        Annual_native = dplyr::coalesce(.data$Annual_native, 0),
        Annual_nonnative = dplyr::coalesce(.data$Annual_nonnative, 0),
        Other_perennial = pmax(0, .data$Cover - .data$Grass - .data$Shrub)
      ) %>%
      dplyr::select(
        Year, Grass, Shrub, Other_perennial, Annual_native, Annual_nonnative, Cover
      )

    stack_long <- stack_ts %>%
      tidyr::pivot_longer(
        cols = c(Grass, Shrub, Other_perennial, Annual_native, Annual_nonnative),
        names_to = "metric",
        values_to = "value"
      ) %>%
      dplyr::mutate(
        # Last factor level sits at bottom (inyoShiny / Lynn): grass → herb → shrub → annuals
        metric = factor(
          metric,
          levels = c(
            "Annual_nonnative", "Annual_native", "Shrub", "Other_perennial", "Grass"
          ),
          labels = c(
            "Annual non-native", "Annual native", "Perennial shrub",
            "Perennial herb", "Perennial grass"
          )
        )
      )

    p_annual_stack <- ggplot(stack_long, aes(x = Year, y = value, fill = metric)) +
      geom_col(position = "stack", width = 0.85, na.rm = TRUE) +
      geom_line(
        data = stack_ts,
        aes(x = Year, y = Cover),
        inherit.aes = FALSE,
        color = "#263238",
        linewidth = 0.7,
        na.rm = TRUE
      ) +
      geom_point(
        data = stack_ts,
        aes(x = Year, y = Cover),
        inherit.aes = FALSE,
        color = "#263238",
        size = 1.2,
        na.rm = TRUE
      ) +
      scale_fill_manual(
        values = c(
          "Perennial grass" = col_grass,
          "Perennial herb" = col_herb,
          "Perennial shrub" = col_shrub,
          "Annual native" = col_ann_nat,
          "Annual non-native" = col_ann_non
        ),
        breaks = c(
          "Perennial grass", "Perennial herb", "Perennial shrub",
          "Annual native", "Annual non-native"
        )
      ) +
      labs(
        x = NULL,
        y = "Cover (%)",
        fill = NULL,
        caption = "Stack = lifeform + annuals; line = total perennial cover (Box test layer)"
      ) +
      scale_x_continuous(breaks = year_breaks, expand = expansion(mult = c(0, 0))) +
      coord_cartesian(xlim = x_limits) +
      panel_theme +
      theme(legend.position = "bottom") +
      guides(fill = guide_legend(nrow = 2, byrow = TRUE))

    if (is.finite(baseline_cover$Cover[1])) {
      p_annual_stack <- p_annual_stack +
        geom_hline(
          yintercept = baseline_cover$Cover[1],
          linetype = "dashed", color = "#263238", linewidth = 0.35
        )
    }
  }

  # 100% stacked perennial lifeform proportions (inyoShiny / Lynn: grass at bottom)
  ts_prop <- ts %>%
    filter(is.finite(pGrass) | is.finite(pShrub) | is.finite(pHerb)) %>%
    mutate(
      Grass = dplyr::coalesce(pGrass, 0),
      Herb = dplyr::coalesce(pHerb, 0),
      Shrub = dplyr::coalesce(pShrub, 0),
      tot = Grass + Herb + Shrub,
      Grass = ifelse(tot > 0, Grass / tot, NA_real_),
      Herb = ifelse(tot > 0, Herb / tot, NA_real_),
      Shrub = ifelse(tot > 0, Shrub / tot, NA_real_)
    ) %>%
    select(Year, Grass, Herb, Shrub) %>%
    pivot_longer(cols = c(Grass, Herb, Shrub), names_to = "metric", values_to = "value") %>%
    # Last factor level at bottom — grass bottom → herb → shrub (same C→B idea as 0.2 grass line)
    mutate(metric = factor(metric, levels = c("Shrub", "Herb", "Grass")))

  # 0.2 grass / shrub≥80% cutpoint lines are Type C→B screens only (not Type B)
  pid_gb <- parcel_gb_type(pid)
  p_stack <- ggplot(ts_prop, aes(x = Year, y = value, fill = metric)) +
    geom_col(position = "stack", width = 0.8, na.rm = TRUE)
  if (type_c_shrub_flag_applies(pid_gb)) {
    p_stack <- p_stack +
      geom_hline(yintercept = 0.2, linetype = "dashed", color = "gray40", linewidth = 0.35) +
      annotate(
        "text", x = year_min, y = 0.22, label = "0.2 grass ref. (Type C→B)",
        hjust = 0, vjust = 0, size = 2.8, color = "gray40"
      )
  } else if (identical(pid_gb, "B")) {
    p_stack <- p_stack +
      annotate(
        "text", x = year_min, y = 0.08, label = "Type B — cutpoint not flagged",
        hjust = 0, vjust = 0, size = 2.6, color = "gray50"
      )
  }
  p_stack <- p_stack +
    scale_fill_manual(
      values = c(Grass = col_grass, Herb = col_herb, Shrub = col_shrub),
      breaks = c("Grass", "Herb", "Shrub"),
      labels = c("Perennial grass", "Perennial herb", "Perennial shrub")
    ) +
    labs(x = NULL, y = "Perennial proportion", fill = NULL) +
    scale_y_continuous(breaks = seq(0, 1, 0.25), limits = c(0, 1), expand = c(0, 0)) +
    scale_x_continuous(breaks = year_breaks, expand = expansion(mult = c(0, 0))) +
    coord_cartesian(xlim = x_limits, clip = "off") +
    panel_theme +
    theme(legend.position = "bottom") +
    guides(fill = guide_legend(nrow = 1))

  # NDVI panel — relative change vs cover (baseline consistency check)
  # Use the parcel's recorded Baseline / Year of baseline measurement.
  ndvi_baseline_year <- profile_baseline_year(pid)
  ndvi_ts <- rs_pfix %>%
    filter(Parcel == pid, Year >= year_min, Year <= year_max) %>%
    group_by(Year) %>%
    summarise(NDVI = round(mean(NDVI_SUR, na.rm = TRUE), 3), .groups = "drop") %>%
    complete(Year = year_min:year_max)

  p_ndvi <- NULL
  if (nrow(ndvi_ts) > 0 && any(is.finite(ndvi_ts$NDVI))) {
    ndvi_baseline <- mean(
      ndvi_ts$NDVI[ndvi_ts$Year == ndvi_baseline_year],
      na.rm = TRUE
    )
    p_ndvi <- ggplot(ndvi_ts, aes(x = Year, y = NDVI)) +
      geom_col(fill = "darkseagreen", width = 0.8, na.rm = TRUE) +
      labs(x = NULL, y = "NDVI (Jul–Sep)") +
      scale_x_continuous(breaks = year_breaks, expand = expansion(mult = c(0, 0))) +
      coord_cartesian(xlim = x_limits) +
      panel_theme
    if (is.finite(ndvi_baseline)) {
      p_ndvi <- p_ndvi +
        geom_hline(yintercept = ndvi_baseline, linetype = "dashed", color = "darkgreen", linewidth = 0.4) +
        annotate(
          "text", x = year_min, y = ndvi_baseline,
          label = paste0(ndvi_baseline_year, " NDVI"),
          hjust = 0, vjust = -0.4, size = 2.8, color = "darkgreen"
        )
    }
  }

  dtw_ts <- dtw_pfix %>%
    filter(Parcel == pid, Year >= year_min, Year <= year_max) %>%
    select(Year, DTW)

  p_dtw <- NULL
  if (nrow(dtw_ts) > 0 && any(!is.na(dtw_ts$DTW))) {
    dtw_stats <- dtw_ts %>%
      group_by(Year) %>%
      summarise(DTW = round(mean(DTW, na.rm = TRUE), 2), .groups = "drop") %>%
      complete(Year = year_min:year_max)

    baseline_avg <- mean(dtw_stats$DTW[dtw_stats$Year == profile_baseline_year(pid)], na.rm = TRUE)

    p_dtw <- ggplot(dtw_stats, aes(x = Year, y = DTW)) +
      geom_line(color = "#2c3e50", linewidth = 0.8, na.rm = TRUE) +
      geom_point(color = "#2c3e50", size = 2.2, na.rm = TRUE) +
      scale_x_continuous(breaks = year_breaks, expand = expansion(mult = c(0, 0))) +
      coord_cartesian(xlim = x_limits) +
      scale_y_reverse(labels = function(x) sprintf("%.2f", x)) +
      labs(x = "Year", y = "DTW (ft BGS)") +
      panel_theme
    if (is.finite(baseline_avg)) {
      p_dtw <- p_dtw +
        geom_hline(yintercept = baseline_avg, linetype = "dashed", color = "#2c3e50") +
        annotate(
          "text", x = year_min, y = baseline_avg,
          label = paste0(profile_baseline_year(pid), " DTW"),
          hjust = 0, vjust = -0.5, size = 3, color = "#2c3e50"
        )
    }
  }

  # Cover first (measurability signal), annual stack, composition, NDVI, DTW
  panels <- list(p_cover)
  heights <- c(1.35)
  if (!is.null(p_annual_stack)) {
    panels <- c(panels, list(p_annual_stack))
    heights <- c(heights, 1.45)
  }
  panels <- c(panels, list(p_stack))
  heights <- c(heights, 1.25)
  if (!is.null(p_ndvi)) {
    panels <- c(panels, list(p_ndvi))
    heights <- c(heights, 1.15)
  }
  if (!is.null(p_dtw)) {
    panels <- c(panels, list(p_dtw))
    heights <- c(heights, 1.15)
  }

  p_combined <- Reduce(`/`, panels) +
    plot_layout(heights = heights, guides = "keep") &
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.box = "horizontal",
      legend.box.just = "left",
      legend.box.spacing = unit(1, "pt"),
      legend.spacing.x = unit(6, "pt"),
      legend.spacing.y = unit(0, "pt"),
      legend.margin = margin(2, 0, 0, 0),
      legend.key.size = unit(0.65, "lines"),
      legend.text = element_text(size = 8),
      plot.margin = margin(4, 16, 4, 8)
    )

  cat("\n\n::: {.parcel-profile-figure}\n\n")
  print(p_combined)
  cat("\n:::\n\n")


  cat('</div>\n')  # parcel-timeseries-block
  cat('</div>\n')  # parcel-profile-main

  # Full-width diagnostics row (was Tufte sticky margin): NMDS + NDVI residual side-by-side
  nmds_png <- here::here("output", "figures", "nmds_parcels", paste0(pid, "_nmds.png"))
  if (!file.exists(nmds_png) && exists("nmds_index") && is.data.frame(nmds_index) && nrow(nmds_index) > 0) {
    hit <- nmds_index %>% dplyr::filter(.data$Parcel == pid, isTRUE(.data$nmds_ok))
    if (nrow(hit) && !is.na(hit$figure[1]) && file.exists(hit$figure[1])) nmds_png <- hit$figure[1]
  }
  ndvi_diag_png <- here::here("output", "figures", "ndvi_residuals", paste0(pid, "_ndvi_residual.png"))

  cat('<aside class="parcel-profile-margin parcel-profile-diagnostics" aria-label="Composition and NDVI diagnostics">\n')
  if (file.exists(nmds_png)) {
    cat('<figure class="parcel-margin-fig">\n')
    cat('<img src="', knitr::image_uri(nmds_png), '" alt="', pid, ' valley-floor NMDS year path with same-TYPE baseline cloud"/>\n', sep = "")
    cat('<figcaption class="parcel-margin-caption"><strong>NMDS path</strong> — valley-floor composition geometry; Type callout = mapped GB type; caption lists same-TYPE cloud n (site-years / parcels); arrows = dominant species centroids + DTW gradient (not control matching).</figcaption>\n')
    cat('</figure>\n')
  }
  if (file.exists(ndvi_diag_png)) {
    cat('<figure class="parcel-margin-fig">\n')
    cat('<img src="', knitr::image_uri(ndvi_diag_png), '" alt="', pid, ' NDVI cover residual diagnostic"/>\n', sep = "")
    cat('<figcaption class="parcel-margin-caption"><strong>NDVI residual</strong> — observed vs expected Cover|NDVI and residual time series; unexpected baseline offset → investigate baseline validity.</figcaption>\n')
    cat('</figure>\n')
  }
  cat('</aside>\n')
  cat('</div>\n')  # parcel-profile-layout

  if (is.null(p_dtw)) {
    cat("**DTW data not available for this parcel — adequacy-for-type cannot be screened from water levels alone.**\n\n")
  }
  if (is.null(p_ndvi)) {
    cat("**NDVI not available for this parcel — skip cover/NDVI consistency check.**\n\n")
  }
}

```
